---
title: "02_Tidy_Tuesday"
author: "Maria Sevillano"
date: "15/01/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tidy Tuesday: Passwords

This R Markdown documents exploration of Tidy Tuesday data found in <https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-01-14>. 

The objective for this week is to visualize popular password and how easy it is to crack down bad passwords.

Since the readme file shows a xkcd comic <https://xkcd.com/936/>, I wanted to incorporate this look into the visualization.  

Here are some resources I found helpful: 

<https://stackoverflow.com/questions/12675147/how-can-we-make-xkcd-style-graphs>

<https://epijim.uk/code-snippets/xkcd/>

## Load libraries
```{r load_libraries, warning=FALSE, message=FALSE}
library(tidyverse)
library(xkcd)
library(RColorBrewer)
library(cowplot)


```


## Get Data
```{r get_data, warning=FALSE, message=FALSE}
passwords <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-14/passwords.csv')

```

## Ispect data
```{r inspect_data, warning=FALSE, message=FALSE}
str(passwords)
summary(passwords)
levels(as.factor(passwords$category))
levels(as.factor(passwords$time_unit))
```

## Wrangle Data
```{r wrangle_data, warning=FALSE, message=FALSE}
font_import(pattern = "[X/x]kcd", prompt=FALSE)
fonttable()
loadfonts()

colors_category<- data.frame(category=levels(as.factor(passwords$category)), color= RColorBrewer::brewer.pal(length(levels(as.factor(passwords$category))), name = "Spectral")) %>%
  mutate(color=as.character(color)) %>%
  column_to_rownames("category")
  

sec_to_day<- function(x){x/(60*60*24)}
min_to_day<- function(x) {x/(60*24)}
hour_to_day<- function(x) {x/24}
week_to_day<- function(x) {7*x}
month_to_day<- function(x) {30.42*x}
year_day<- function(x) {365*x}


for (i in levels(as.factor(passwords$category))){
  #Data for left side
  passwords1<- passwords %>%
    filter(category== i) %>%
    group_by(category) %>%
    top_n(-20, rank) %>%
    mutate(pass1= as.numeric(as.factor(password))) %>%
    mutate(xmin= pass1 - 0.1) %>%
    mutate(xmax= pass1 + 0.1) %>%
    mutate(ymin = 0) %>%
    mutate(ymax = offline_crack_sec)
  
  #Give some space to fit in text
  xrange1 <- range(min(passwords1$xmin), max(passwords1$xmax))
  yrange1 <- range(min(passwords1$ymin), max(passwords1$ymax))
  mapping1 <- aes(xmin=xmin,ymin=ymin,xmax=xmax,ymax=ymax, fill=category)
  
  #Data for right side
  passwords2<- passwords %>%
    filter(category== i) %>%
    group_by(category) %>%
    top_n(-20, rank) %>%
    mutate(pass2= as.numeric(as.factor(password))) %>%
    mutate(value2= ifelse(time_unit=="hours", hour_to_day(value), 
                         ifelse(time_unit=="minutes", min_to_day(value),
                                  ifelse(time_unit=="months", month_to_day(value),
                                         ifelse(time_unit=="seconds", sec_to_day(value),
                                                ifelse(time_unit=="weeks", week_to_day(value),
                                                       ifelse(time_unit=="years", year_day(value),
                                                              ifelse(time_unit=="days", value, value)))))))) %>%
    mutate(xmin= pass2 - 0.1) %>%
    mutate(xmax= pass2 + 0.1) %>%
    mutate(ymin = 0) %>%
    mutate(ymax = value2)
  
  #Give some space to fit in text
  xrange2 <- range(min(passwords2$xmin), max(passwords2$xmax))
  yrange2 <- range(min(passwords2$ymin), max(passwords2$ymax))
  mapping2 <- aes(xmin=xmin,ymin=ymin,xmax=xmax,ymax=ymax, fill=category)
  
  
  #Plots
  p1<- ggplot() + 
    xkcdrect(mapping1, passwords1) +
    xkcdaxis(xrange1,yrange1) +
    scale_fill_manual(values = colors_category[i,"color"]) +
    scale_y_reverse() +
    coord_flip()  +
    ggtitle(i, subtitle = "Crack offline") +
    theme(axis.title.y = element_blank(), 
          axis.text.y = element_blank(), 
          axis.ticks.y = element_blank(),
          legend.position = "none",
          plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5)) +
    ylab("Time, in sec")
  
  
  pmid<- ggplot(passwords1, aes(x=0,y=pass1)) +
    geom_text(aes(label=password, size= font_size, family="xkcd"), color= colors_category[i,"color"]) +
    xkcdaxis(c(0,0),xrange1) +
    ggtitle("Popularity") +
    theme_xkcd()+
    theme(axis.title=element_blank(),
          panel.grid=element_blank(),
          axis.text.y=element_blank(),
          axis.ticks.y=element_blank(),
          panel.background=element_blank(),
          axis.text.x=element_text(color=NA),
          axis.ticks.x=element_line(color=NA),
          legend.position = "none",
          plot.title = element_text(hjust = 0.5)) 
  
  p2<- ggplot() + 
    xkcdrect(mapping2, passwords2) +
    xkcdaxis(xrange2,yrange2) +
    scale_fill_manual(values = colors_category[i,"color"]) +
    coord_flip()  +
    ggtitle(i, subtitle = "Crack manually") +
    theme(axis.title.y = element_blank(), 
          axis.text.y = element_blank(), 
          axis.ticks.y = element_blank(),
          legend.position = "none",
          plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5)) +
    ylab("Time, in days")
  
  plot<- plot_grid(p1,pmid, p2,align = "h", nrow = 1, axis = "tblr", rel_widths = c(1,0.5,1))
  print(plot)
}


```

# Alternative #1
You cant't really see a difference in bar sizes when using year as time unit. 
```{r alternative1, warning=FALSE, message=FALSE}

for (i in levels(as.factor(passwords$category))){
  #Data for left side
  passwords1<- passwords %>%
    filter(category== i) %>%
    group_by(category) %>%
    top_n(-20, rank) %>%
    mutate(pass1= as.numeric(as.factor(password))) %>%
    mutate(xmin= pass1 - 0.1) %>%
    mutate(xmax= pass1 + 0.1) %>%
    mutate(ymin = 0) %>%
    mutate(ymax = offline_crack_sec)
  
  #Give some space to fit in text
  xrange1 <- range(min(passwords1$xmin), max(passwords1$xmax))
  yrange1 <- range(min(passwords1$ymin), max(passwords1$ymax))
  mapping1 <- aes(xmin=xmin,ymin=ymin,xmax=xmax,ymax=ymax, fill=category)
  
  #Data for right side
  passwords2<- passwords %>%
    filter(category== i) %>%
    group_by(category) %>%
    top_n(-20, rank) %>%
    mutate(pass2= as.numeric(as.factor(password))) %>%
    mutate(xmin= pass2 - 0.1) %>%
    mutate(xmax= pass2 + 0.1) %>%
    mutate(ymin = 0) %>%
    mutate(ymax = value)
  
  #Give some space to fit in text
  xrange2 <- range(min(passwords2$xmin), max(passwords2$xmax))
  yrange2 <- range(min(passwords2$ymin), max(passwords2$ymax))
  mapping2 <- aes(xmin=xmin,ymin=ymin,xmax=xmax,ymax=ymax, fill=category)
  
  
  #Plots
  p1<- ggplot() + 
    xkcdrect(mapping1, passwords1) +
    xkcdaxis(xrange1,yrange1)  +
    scale_y_reverse() +
    scale_fill_manual(values = colors_category[i,"color"]) +
    coord_flip()  +
    ggtitle(i, subtitle = "Crack offline") +
    theme(axis.title.y = element_blank(), 
          axis.text.y = element_blank(), 
          axis.ticks.y = element_blank(),
          legend.position = "none",
          plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5)) +
    ylab("Time, sec")
  
  
  pmid<- ggplot(passwords1, aes(x=0,y=pass1)) +
    geom_text(aes(label=password, size= font_size, family="xkcd"), colour= colors_category[i,"color"]) +
    xkcdaxis(c(0,0),xrange1, colour=colors_category[i,"color"]) +
    ggtitle("Popularity") +
    theme_xkcd()+
    theme(axis.title=element_blank(),
          panel.grid=element_blank(),
          axis.text.y=element_blank(),
          axis.ticks.y=element_blank(),
          panel.background=element_blank(),
          axis.text.x=element_text(color=NA),
          axis.ticks.x=element_line(color=NA),
          legend.position = "none",
          plot.title = element_text(hjust = 0.5)) 
  
  p2<- ggplot() + 
    xkcdrect(mapping2, passwords2) +
    xkcdaxis(xrange2,yrange2) +
    geom_text(data = passwords2, aes(x= pass2, y= value + 0.5, label=time_unit, family="xkcd") , color= colors_category[i,"color"], hjust=1, fill= "white") +
    scale_fill_manual(values = colors_category[i,"color"]) +
    coord_flip()  +
    ggtitle(i, subtitle = "Crack manually") +
    theme(axis.title.y = element_blank(), 
          axis.text.y = element_blank(), 
          axis.ticks.y = element_blank(),
          legend.position = "none",
          plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5)) +
    ylab("Time")
  
  plot<- plot_grid(p1,pmid, p2,align = "h", nrow = 1, axis = "tblr", rel_widths = c(1,0.5,1))
  print(plot)
}
```

# Alternative #2
```{r alternative2, warning=FALSE, message=FALSE}
password4<- passwords %>%
  filter(category== "cool-macho") %>%
  group_by(category) %>%
  top_n(-20, rank) %>%
  ungroup() %>%
  mutate(offline_crack_sec= -offline_crack_sec) %>%
  select(password, value, offline_crack_sec) %>%
  gather(key="crack", value= "time", -password)


password5<- passwords %>%
  select(password, rank)

password4 <- password4 %>%
  left_join(password5)

breaks_values <- pretty(password4$time)

ggplot(password4) +
  geom_bar(aes(reorder(password, desc(rank)), time, fill=crack), stat = "identity") +
  scale_y_continuous(breaks = breaks_values,
                     labels = abs(breaks_values)) +
  coord_flip() +
  xlab("Password") +
  theme_xkcd()

```

