---
title: "04_TidyTuesday"
author: "Maria Sevillano"
date: "02/02/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tidy Tuesday: San Francisco Trees

This R Markdown documents exploration of [Tidy Tuesday](https://github.com/rfordatascience/tidytuesday/tree/master/data/2020/2020-01-28).  

The objective for this week is to:  

* visualize how the number and species of trees has changed over the years since 1955.  

These are good resources to learn about maps:  
1. <https://dadascience.design/post/r-gis-like-it-s-2019-pragmatic-workflows-pt-1/>
2. <https://bhaskarvk.github.io/user2017.geodataviz/notebooks/02-Static-Maps.nb.html#animations>

Here's a [link](https://medium.com/@SportSciData/https-medium-com-collinsneil306-how-to-create-interactive-reports-with-r-markdown-part-i-4fa9df46cd9) for data tables.
## Load libraries
```{r load_libraries, warning=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
library(gganimate)
library(sf)
library("rnaturalearth")
library("rnaturalearthdata")
library(magick)
#library(DT)
library(knitr)
library(kableExtra)

```

## Get Data
```{r get_data, warning=FALSE, message=FALSE}
sf_trees <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-28/sf_trees.csv')

```

For the species column;
'Tree(s) ::' label doesn't provide additional information about the tree type, will filter out.
'::' label doesn't provide additional information about the tree type, will filter out

The '::' delimiter separates the scientific name and common name of tree. 
Date Planted (NA if before 1955), will only keep those planted after 1955.


## Check Data Structure
```{r data_structure, warning=FALSE, message=FALSE}
str(sf_trees)
summary(sf_trees)

levels(as.factor(sf_trees$legal_status))
levels(as.factor(sf_trees$species))
levels(as.factor(sf_trees$caretaker))

```

## Wrangle data
```{r data_wrangling, warning=FALSE, message=FALSE}

sf_trees_tidy <- sf_trees %>%
  filter(species != "::") %>%
  filter(species != "Tree(s) ::") %>%
  filter(!is.na(date)) #These filtering steps remove ~ 70% of the data, but we still have 58090 observations.


length(levels(as.factor(sf_trees_tidy$species))) 

sf_trees_tidy <- sf_trees_tidy %>%
  separate(species, into = c("scientific", "common"), sep = "::") %>%
  separate(scientific, into= c("genus", "species"), sep= " ") %>%
  mutate(year_planted= year(date)) %>%
  mutate(keeper= ifelse(caretaker=="Private", "Private", "Other"))

sf_trees_tidy %>% group_by(keeper) %>% summarise(counts= n())

selected_genus<- sf_trees_tidy %>% 
  group_by(genus, species) %>%
  summarise(counts= n()) %>%
  ungroup() %>%
  spread(key=genus, value=counts) %>%
  select(everything()) %>%  # replace to your needs
  summarise_all(funs(sum(!is.na(.)))) %>%
  select(-species) %>%
  gather(key = "genus", value = "richness") %>%
  arrange(desc(richness)) %>%
  as.data.frame()

#DT::datatable(selected_genus, class = 'cell-border stripe', rownames = F, filter = 'top', options = list(pageLength = 10), width = '55%')

kable(selected_genus) %>%
  kable_styling() %>%
  scroll_box(width = "500px", height = "200px")

#to try out plot
random_subset<- sample_n(sf_trees_tidy, 200)

selected_sf_trees<- sf_trees_tidy %>%
  filter(genus== "Melaleuca") %>%
  mutate(common=as.factor(common))
 
```

## Plot
A big reccommendation is to keep track of package version when troubleshooting or following tutorials.  

[This](http://www.bernhardlearns.com/2018/11/animating-r-plot-with-gganimate.html) tutorial was useful in deciphering how to add points in every transition state, without removing ponts past transition states.

Below is an annimation of all the records from 1955 that have a tree species designation. 
```{r plots, warning=FALSE, message=FALSE}

california_map <- ne_states(country = "united states of america", returnclass = "sf") %>%
  filter(postal == "CA")

class(california_map)

#http://www.bernhardlearns.com/2018/11/animating-r-plot-with-gganimate.html
p<- ggplot() +
  geom_sf(data = california_map, fill= "antiquewhite") +
  coord_sf(xlim = c(-122.3, -122.6), ylim = c(37.7, 37.84)) +
  geom_point(data= sf_trees_tidy, aes(longitude, latitude, group=year_planted, shape= keeper), color= "#006700", alpha=0.5) +
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(), 
        panel.background = element_rect(fill = "lightblue"),
        panel.grid.major = element_line(colour = "lightblue"), 
        panel.grid.minor = element_line(colour = "lightblue")) +
  transition_states(year_planted, 3,1) +
  shadow_mark(past=TRUE) +
  ggtitle('Now showing year {closest_state}')

p
```

An approach to reduce dimensions of the data is to subset genera of interest.

```{r plots2, warning=FALSE, message=FALSE}
p2<- ggplot() +
  geom_sf(data = california_map, fill= "antiquewhite") +
  coord_sf(xlim = c(-122.3, -122.6), ylim = c(37.7, 37.84)) +
  geom_point(data= selected_sf_trees, aes(longitude, latitude, group=year_planted, color= common), shape= 18, alpha=0.5) +
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(), 
        panel.background = element_rect(fill = "lightblue"),
        panel.grid.major = element_line(colour = "lightblue"), 
        panel.grid.minor = element_line(colour = "lightblue")) +
  transition_states(year_planted, transition_length = 3, state_length =1) +
  shadow_mark(past=TRUE) +
  ggtitle('Now showing year {closest_state}') +
  theme(legend.position="bottom")  

p2
```

Randomly zoomed to an area
```{r plot3, warning=FALSE, message=FALSE}

p3<- ggplot() +
  geom_sf(data = california_map, fill= "antiquewhite") +
  coord_sf(xlim = c(-122.3, -122.6), ylim = c(37.7, 37.84)) +
  geom_point(data= selected_sf_trees, aes(longitude, latitude, group=year_planted, color= common), shape= 18, alpha=0.5) +
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(), 
        panel.background = element_rect(fill = "lightblue"),
        panel.grid.major = element_line(colour = "lightblue"), 
        panel.grid.minor = element_line(colour = "lightblue")) +
  transition_states(year_planted, 3,1) +
  shadow_mark(past=TRUE) +
  ggtitle('Now showing year {closest_state}') +
  theme(legend.position="bottom") + 
  view_step(fixed_x = c(-122.45, -122.4), fixed_y = c(37.78, 37.8), pause_length = 1, step_length = 2, nsteps = 3)
p3

```


