---
title: "Tidy_Tuesday_2020_04_14"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tidy Tuesday: Tour de France

The objective for this week is to analyse data about Tour de France. We often forget that tables are very useful visualization and prefered in [certain situations](https://flourish.studio/2019/03/11/how-make-interactive-table/). Because of this, I decided to give the package `gt` a try. This package allows you to showcase *great looking tables* using gramar of tables much like ggplot2 grammar of graphics. You can find an introductury tutorial [here](https://blog.rstudio.com/2020/04/08/great-looking-tables-gt-0-2/).


Back to traditional visualization, I would like to see the effect of the numeric variables (age, height, weight, distance) on the overall time (time_overall) to complete the race. 

Also of interest is the age of the winner, assuming the won at peak athletic performance and the age the participant died at.

margin between winner and runner up

## Load libraries

```{r load_libraries, warning=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
library(gt)
library(skimr)
library(ggthemes)
```

## Get Data

```{r get_data, warning=FALSE, message=FALSE}
tdf_winners <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-04-07/tdf_winners.csv')

```

## Data structure
```{r data_structure}
skim(tdf_winners)

```

## Data Wrangling 

```{r data_wrangling, message=FALSE}
#gt

tdf_winners %>%
  select(winner_name, birth_country, birth_town) %>%
  distinct %>%
  group_by(birth_country, birth_town) %>%
  summarize(Winners= n()) %>%
  gt(rowname_col = "birth_town") %>%
  tab_header(title = md("&#128690; Tour de France &#128690;")) %>%
  summary_rows(
    groups = TRUE,
    columns = vars(Winners),
    fns = list(TOTAL = "sum"),
    formatter = fmt_number,
    decimals = 0,
    use_seps = TRUE
  ) %>%
  opt_all_caps(locations = "row_group") %>%
  tab_options(
    summary_row.background.color = "black",
    row_group.background.color = "yellow",
    heading.background.color = "white",
    heading.title.font.size = 20,
    column_labels.background.color = "black",
    stub.background.color = "white",
    stub.border.style = "dashed",
    stub.border.width = "1px",
    summary_row.border.color = "#989898",
    table.width = "50%"
  )

tdf_time<- tdf_winners %>%
  select(age, height, weight, distance, time_overall) %>%
  drop_na() %>%
  gather(variable, value, -time_overall)

tdf_margin<- tdf_winners %>%
  mutate(overall_second= time_overall + time_margin, year= year(start_date)) %>%
  select(year, time_overall, overall_second) %>%
  drop_na()


```

## Visualizations

```{r }
ggplot(tdf_time) +
  theme_few() +
  geom_point(aes(value, time_overall)) +
  facet_wrap(~variable, scales = "free") +
  ylab("Overall time in race (hrs)") +
  ggtitle("Tour de France winners ") +
  theme(axis.title.x = element_blank()) +
  theme(panel.background = element_rect(fill="yellow")) 

ggplot(tdf_margin) +
  theme_few() +
  geom_segment(aes(x=overall_second, xend=time_overall, y= year, yend=year), size=5, color="yellow") +
  scale_y_continuous(limits = c(1900, 2020), breaks = seq(1900, 2020, by = 5)) +
  ylab("Edition") +
  xlab("Overall time in race (hrs)") +
  theme(panel.background = element_rect(fill="black")) +
  geom_text(aes(200, 2010), label="The margin between runner up and winner has \nshrunk over the years", color= "white")

```


