---
title: "Tidy_Tuesday_2020_05_12 "
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tidy Tuesday: Animal Crossing

The objective this week is to visualize [data](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-05-05/readme.md){target="_blank"} from Animal Crossing, a video game released by Nintendo. Particularly, I want to visualize the distribution of character personalities by different species/gender. Additionally, I will represent real evolutionary distances between top species with a phylogenetic tree with tips labeled as character of that particular species. The newick tree was generated using [phyloTree](https://phylot.biobyte.de/index.cgi){target="_blank"}.


## Load libraries
```{r load_libraries, warning=FALSE, message=FALSE}
library(tidyverse)
library(readxl)
library(ggimage)
library(ggtree)
library(ggtext)
library(gcookbook)
library(hrbrthemes)
library(extrafont)
library(patchwork)
library(gt)

```

## Get Data
```{r get_data, warning=FALSE, message=FALSE}
items <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-05/items.csv')
villagers <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-05/villagers.csv')
mapping_species<- read_excel("species_mapping.xlsx", col_names = c("tip_label", "species"))
phyloT<- read.tree("phyloT.nwk")

```

## Data Wrangling 
```{r data_wrangling, message=FALSE}
#font_import()
#fonts()

villagers_species_gender_personality<- villagers %>%
  group_by(species, gender, personality) %>%
  tally %>%
  filter(!personality=="uchi")


top_5_species<- villagers %>% 
  group_by(species) %>% 
  tally %>% 
  arrange(desc(n)) %>% 
  top_n(5) %>%
  pull(species)

mapping_info<- villagers %>%
  filter(species %in% top_5_species) %>%
  filter(personality=="normal") %>%
  group_by(species) %>%
  filter(row_number()==1) %>%
  select(url) %>%
  left_join(mapping_species) %>%
  mutate(file_name= str_extract(url, "(?<=thumb/).+"))


tree_tip<- as.data.frame(phyloT$tip.label) 
colnames(tree_tip)<- "NCBI"
tree_tip<- tree_tip %>%
  mutate(name= plyr::mapvalues(NCBI, mapping_info$tip_label, mapping_info$file_name))

phyloT$tip.label<- as.character(tree_tip$name)
```


## Visualizations

Table format
```{r fig.width= 10, fig.height=8, warning=FALSE, message=FALSE}

villagers_species_gender_personality %>%
  filter(species %in% top_5_species) %>% 
  ungroup() %>%
  pivot_wider(names_from = species, values_from = n, values_fill = list(n=0)) %>%
  group_by(gender) %>%
  gt(rowname_col = "personality") %>%
  tab_spanner(
    label = "Species",
    columns = vars(cat, duck, frog, rabbit, squirrel)
    ) 
```

Heatmap and phylogenetic tree.
The character images were chosen quasi randomly. 

```{r fig.width= 10, fig.height=8, warning=FALSE, message=FALSE}
 
hm<- ggplot(villagers_species_gender_personality %>% filter(species %in% top_5_species) , aes(species, personality)) +
  geom_tile(aes(fill=n)) +
  facet_grid(~gender, scale="free") +
  scale_fill_viridis_c(option = "magma") +
  theme_ipsum(grid="XY", axis="xy") +
  theme(legend.position = "none") +
   labs(
    title="Number of species in Animal Crossing",
    subtitle="Only top 5 species are shown \nDarker colors represent fewer observations",
    caption="Source: VillagerDB"
  ) 

pT<- ggtree(phyloT, layout="circular") + 
  geom_tiplab(aes(image=paste0("images/", label)), 
                geom="image")

hm + pT

```
