---
title: "Tidy_Tuesday_2020_04_21 "
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tidy Tuesday: Best Rap Artists

This week's objective is to visualize [data](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-04-14/readme.md) on hip hop songs rankings.  

I came across [this](https://resources.rstudio.com/rstudio-conf-2020/the-glamour-of-graphics-william-chase?wvideo=fhb2ifduim) presentation by [William Chase](https://www.williamrchase.com/) on the Glamour of Graphics. The main takeaway is that in order to make aesthetically pleasing visualizations you should pay attention to 3 characteristics: (1) layout, (2) typography, and (3) color. 


1. **Layout**   
    * Appropriate use of *titles* can go a long way in decluttering your plots, it makes axis titles redundant and could aid with legend positioning or removal. See `ggtext` [package](https://github.com/wilkelab/ggtext)    
    * *Alignments* create symmetry and a seamless experience for redability. Align titles all the way to the start of the plot (including axis text). Avoid having people tilt their heads: rotate your plot! 
    * Mind your lines, question if background *grid lines* are neccessary.  
    * Don't be afraid of *whitespace*: less is more! If you cram in too much information, the message can get lost. 
    
2. **Typography**
    * Make use of other *fonts*, more often than not built-in fonts are outdated. Download [free fonts](fonts.google.com)
    * Maintain *hierarchy* (size, weight, color, spacing, contrasting typespace) to guide the reader. 
    * Avoid using oldstyle *numbering*, do use lining numbering in which the height is consistent. Similarly, avoid using proportional numbering, do use a tabular kind for width consistency.  

3. **Color**
    * Think about *color theory*. When choosing hues, use color wheel to convey sentiments: 
          - complementary (high contrast)  
          - analogous (calm, harmonious)  
          - triadic (vibrant, contrast)  
    * Control colors by Hue, saturation, lightness/brightness (*HSB*).
    * Create *your own palette* using an eyedropper tool (e.g. color slurp tool for Mac)
    * Mind *accessibility*, think about color-blindness for example.
    * For plot *backgrounds*, avoid completely white or black, pastel colors are a good option. 


Using these principles I would like to see how songs get ranked across time, and which songs are ranked the highest. 

## Load libraries

```{r load_libraries, warning=FALSE, message=FALSE}
library(tidyverse)
library(skimr)
library(patchwork)
library(ggpubr)
library(cowplot)
library(ggtext)

```

## Get Data

```{r get_data, warning=FALSE, message=FALSE}
polls <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-04-14/polls.csv')
rankings <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-04-14/rankings.csv')

```

## Data structure
```{r data_structure}
skim(polls)
skim(rankings)
```

## Data Wrangling 

```{r data_wrangling, message=FALSE}
rank_freq<- rankings %>% 
  group_by(year) %>% 
  summarise(Freq= n())

rank_top<- rankings %>% 
  group_by(gender) %>% 
  top_n(1, points) %>% 
  mutate(label= paste(title, artist, year, sep = "\n"))

```

## Visualizations

```{r fig.width= 10, fig.height=8}
p1<-ggplot() +
  theme_minimal() +
  geom_bar(data= rank_freq, aes(year, Freq), stat = "identity") +
  geom_bar(data= rank_freq %>% filter(year>=1990 & year<=2000), aes(year, Freq), stat = "identity", fill = "gold") +
  geom_vline(xintercept = 1990) +
  geom_vline(xintercept = 2000) +
  geom_bracket(xmin = 1990, xmax = 2000, y.position = 30,
    label = "The golden era of Hip-hop") +
  ggtitle("Number of songs ranked for each year: the 90's don't disappoint") +
  theme(plot.title.position = "plot", axis.title = element_blank(), panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank())


p2<- ggplot() +
  theme_minimal() +
  geom_point(data= rankings, aes(year, points, color=gender), alpha=0.8, size=2)  +
  geom_vline(xintercept = 1990) +
  geom_vline(xintercept = 2000) +
  theme(legend.position = "none") +
  ylab("Points") +
  geom_point(data= rankings, aes(year, points, color=gender), alpha=0.8, size=2, shape= 1, color="black") +
  geom_curve(data= rank_top, aes(xend=year , yend= points, x = c(2012, 2005, 2010), y=points + 20),
             curvature = .2, arrow = arrow(length = unit(1, "mm")), color="gold") +
  geom_label(data=rank_top, aes(x=c(2012, 2005, 2010), points + 20, label=label, fill=gender), color= "gold", fontface = "bold") +
  scale_color_manual(values=c("grey", "black", "white")) +
  scale_fill_manual(values=c("grey", "black", "white")) +
  labs(title = "Number of points for songs by <strong>male</strong></b>, 
    <strong><span style='color:#808080'>female</span></strong></b>, and <strong><span style='color:#FFFFFF'>mixed artists</span></strong></b> by year") +
  theme(plot.title = element_textbox_simple(
      size = 13,
      lineheight = 1,
      padding = margin(5.5, 5.5, 5.5, 5.5),
      margin = margin(0, 0, 5.5, 0),
      fill = "gold"
    ), 
        plot.title.position = "plot", axis.title = element_blank(), panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank())

p1
p2

```


